Collaborative Visual Analysis on the Web with RCloud

### 2015-06-26 system.tex
copy editing

Our design takes advantage of existing web software as much as possible.
HTTP, despite its deficiencies, is the lingua franca of distributed
interprocess communication. By speaking HTTP natively, our system
provides a low-friction path from experimental data analysis scripts
to automated \emph{web services}. 
essentially a remote function call over
the web, which can be invoked by higher-level tools. An example of
this sort of progression is described in Section~\ref{sec:casestudy}.

@@

RCloud's implementation of the versioning mechanism is built on
top of GitHub's \emph{gists}~\cite{GitHub:2014:GG}, which are an HTTP
interface for simplified repositories. The GitHub web-service API
provides most of the semantics we need for the versioning portion of
the storage back end: access to previous versions, comments, starring,
and forking. This provides the additional benefit that every RCloud
notebook can also be manipulated like a git repository. Advanced users
then have access to features like command-line checkout, history
editing, and so on.

@@

The same idea of exposing functionality via hashes can be used to give
the server-side of RCloud access to Javascript functions. This allows
R libraries to respond to user input in the browser, giving them
access to features ranging from password prompts, to the currently
selected set of points in an interactive linked brushing
visualization.

As a result, the features required to provide safe access to R from
the client, and Javascript access from the R side, also enable
full two-way communication between the languages.  This provides
considerable flexibility, so that for example, a chart built with dc.js
or leaflet.js can call back to analysis functions in R without having
to define an additional protocol between the processes. From the
Javascript side, an RPC call into the R process is just another HTTP
request. From the R process, a call into Javascript looks like just
another function call.

>>>>

Our design takes advantage of existing web software as much as possible.
HTTP, despite its deficiencies, is the lingua franca of distributed
interprocess communication. By speaking HTTP natively, our system
provides a low-friction path from experimental data analysis scripts
to automated \emph{web services}: essentially a remote function call over
the web which can be invoked by higher-level tools.
%An example of
%this sort of progression is described in Section~\ref{sec:casestudy}.

@@

RCloud's implementation of the versioning mechanism is built on
top of GitHub's \emph{gists}~\cite{GitHub:2014:GG}, which are an HTTP
interface for simplified repositories. The GitHub web-service API
provides most of the semantics we need for the versioning portion of
the storage back end: access to previous versions, comments, starring,
and forking. This provides the additional benefit that every RCloud
notebook can also be manipulated like a git repository, granting advanced users
access to features like command-line checkout and history
editing.

@@

The same idea of exposing functionality via hashes can be used to give
the server-side of RCloud access to Javascript functions. This allows
R libraries to request user input in the browser, giving them
access to features ranging from password prompts, to the currently
selected set of points in an interactive visualization.

As a result, the features required to provide safe access to R from
the client, and Javascript access from the R side, also enable
full two-way communication between the languages.  This provides
considerable flexibility, so that for example, a chart built with dc.js
or leaflet.js can call analysis functions in R without having
to define an additional protocol between the processes. Calls in either
direction look like ordinary function calls.
%% gw: inaccurate, it's websocket; we still don't support HTTP into a live process
%% From the
%% Javascript side, an RPC call into the R process is just another HTTP
%% request. From the R process, a call into Javascript looks like just
%% another function call.
